Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-21 paper.tex
remv whitespaces padding $$

\newcommand{\typingHC}[3]{#1 : #2 \Longrightarrow #3}
\newcommand{\hcvOOinj}[2]{\mathtt{inj} \; #2}
\newcommand{\hcvOOfun}[5]{\mathtt{fun} \; #1 \; #2 \; #3 \; #4 \; #5}
\newcommand{\hcvOOtt}[0]{\mathtt{unit}}
\newcommand{\hcvOOcons}[4]{\mathtt{cons}\;#1\;#2\;#3\;#4}
\newcommand{\hcvOOinl}[2]{\mathtt{inl}\;#1\;#2}
\newcommand{\hcvOOinr}[2]{\mathtt{inr}\;#1\;#2}
\newcommand{\hckOOmt}[0]{\mathtt{stop}}
\newcommand{\hckOOconsI}[3]{\mathtt{cons_1}\;#1\;#2\;#3}
\newcommand{\hckOOappII}[2]{\mathtt{app_2}\;#1\;#2}
\newcommand{\sidecond}[1]{\text{if}\;#1}
% Lazy D cast calculus on space-inefficient CEK
\newcommand{\judgeCreduce}[2]{#1 \longmapsto_{\mathcal{C}} #2}
\newcommand{\judgeCreduceTrans}[2]{#1 \longmapsto_{\mathcal{C}}^{*} #2}
\newcommand{\judgeCeval}[2]{eval_{\mathcal{C}}(#1) = #2}
\newcommand{\redrule}[3]{#1 & \longmapsto_\mathcal{C} & #2 & #3\\}
\newcommand{\hiredrule}[3]{\highlight{#1} & \highlight{\longmapsto_\mathcal{C}} & \highlight{#2} & \highlight{#3} \\}
% blame calculus on space-efficient CEK
\newcommand{\judgeSreduce}[3]{#2 \longmapsto_{\mathcal{S}(#1)} #3}
\newcommand{\judgeSreduceTrans}[3]{#2 \longmapsto_{\mathcal{S}(#1)}^{*} #3}
\newcommand{\judgeSeval}[3]{eval_{\mathcal{S}(#1)}(#2) = #3}
\newcommand{\redruleS}[3]{#1 & \longmapsto_{\mathcal{S}(C)} & #2 & #3\\}
\newcommand{\hiredruleS}[3]{\highlight{#1} & 
\highlight{\longmapsto_{\mathcal{S}(C)}} & \highlight{#2} & \highlight{#3} \\}
% Normal Coercion
\newcommand{\ncProj}[2]{#1?^{#2}}
\newcommand{\ncInj}[1]{#1!}
\newcommand{\ncId}[0]{\iota}
\newcommand{\ncSeq}[2]{#1;#2}
\newcommand{\ncFail}[1]{\bot^{#1}}
\newcommand{\ncFun}[2]{\POOfun{#1}{#2}}
\newcommand{\ncProd}[2]{\POOprod{#1}{#2}}
\newcommand{\ncSum}[2]{\POOsum{#1}{#2}}
% Hypercoercion
\newcommand{\hyperCoercionI}[0]{\mathtt{id\star}}
\newcommand{\hyperCoercionC}[3]{#1 \overset{#2}{\curvearrowright} #3}
% machine state simulations
\newcommand{\eqvS}[4]{#3 \approx_{\mathcal{S}\mathcal{S}} #4}
\newcommand{\eqvSD}[3]{#2 \approx_{\mathcal{SD}} #3}
% to-dos
\newcommand{\todoKC}[1]{\todo[inline]{KC needs to #1}}
\newcommand{\todoKCFixed}[0]{\todo[inline]{Fixed. -KC}}
% abbrev
\newcommand{\castCalculus}[0]{$\lambda_{\rightarrow}^{\langle\cdot\rangle}$}
% names
\newcommand{\ineffCEK}{$ \mathcal{C} $}
\newcommand{\ineffCEKD}{$ \mathcal{D} $}
\newcommand{\ineffCEKUD}{$ \mathcal{UD} $}
\newcommand{\judgeDreduce}[2]{#1 \longmapsto_{\mathcal{D}} #2}
\newcommand{\judgeDreduceTrans}[2]{#1 \longmapsto_{\mathcal{D}}^{*} #2}
\newcommand{\judgeDeval}[2]{eval_{\mathcal{D}}(#1) = #2}
\newcommand{\judgeUDreduce}[2]{#1 \longmapsto_{\mathcal{UD}} #2}
\newcommand{\judgeUDreduceTrans}[2]{#1 \longmapsto_{\mathcal{UD}}^{*} #2}
\newcommand{\judgeUDeval}[2]{eval_{\mathcal{UD}}(#1) = #2}
\newcommand{\effCEK}[1]{$ \mathcal{S}(#1) $}
\newcommand{\evalEqv}[2]{$ eval_{\text{#1}} = eval_{\text{#2}} $}
\newcommand{\continue}[2]{continue(#2,#1)}
\newcommand{\highlight}[1]{{\color{red} #1}}

@@

	Consistency
	\fbox{$ T_1 \sim T_2 $}
	\begin{gather*}
	\inference{}{
		\star \sim \star
	} \quad
	\inference{}{
		\star \sim P
	} \quad
	\inference{}{
		P \sim \star
	} \\[1ex]
	\inference{}{
		\iota \sim \iota
	} \quad
	\inference{
		S_1 \sim S_2 &
		T_1 \sim T_2
	}{
		S_1 \rightarrow T_1 \sim S_2 \rightarrow T_2
	} \quad
	\inference{
		S_1 \sim S_2 &
		T_1 \sim T_2
	}{
		S_1 \times T_1 \sim S_2 \times T_2
	} \quad
	\inference{
		S_1 \sim S_2 &
		T_1 \sim T_2
	}{
		S_1 \plus T_1 \sim S_2 \plus T_2
	}
	\end{gather*}

	Term typing
	\fbox{$ \judgetype{\Gamma}{e}{T} $}
	\begin{gather*}
          \dots \qquad
		\inference{
			\Gamma \vdash e : T_1 & T_1 \sim T_2
		}{
			\judgetype{\Gamma}{\eOOcast{e}{T_1}{l}{T_2}}{T_2}
		} \quad
		\inference{
		}{
			\judgetype{\Gamma}{\eOOblame{l}}{T}
		}
	\end{gather*}

@@

The syntax and static semantics is the same for the \lazyD{} and
\lazyUD{} cast calculi, and is reviewed in \figref{fig:blame-static}.
As usual, the important features are the cast expressions,
$\eOOcast{e}{T_1}{l}{T_2}$, which are responsible for runtime type
checking, and blame expressions, $\eOOblame{l}$, that raise errors.
The syntax and static semantics is the same that of
\citet{siek2009exploring} except for a few minor exceptions:
\begin{itemize}
\item We add sum, product, and unit types.
\item We separate types into those with a type constructor at the top,
  the \emph{pretypes} ($\POOunit$, functions, products, and sums),
  versus the dynamic type $\star$ (a.k.a. $ \mathtt{Dyn} $ or
  $\mathbb{?}$).
\item We annotate the codomain of a lambda abstraction explicitly
  because we refer to it in the dynamic semantics. 
\end{itemize}
As usual, the source $T_1$ and target types $T_2$ of a cast
$\eOOcast{e}{T_1}{l}{T_2}$ are required to be consistent, written $T_1
\sim T_2$. The consistency relation is standard and defined in
\figref{fig:blame-static}.

%% defines the syntax of the cast calculus and
%% its static semantics. We extend the syntax of
%%  with sum and product types.  Let $ T $ range
%% over types. A type is either the dynamic type $ \star $ (a.k.a. $
%% \mathtt{Dyn} $, $ \mathbb{?} $, or $ \mathtt{Unknown} $), or a type

%% As usual we write $ T_1 \sim T_2 $ when $ T_1 $ and $ T_2 $ are
%% consistent. The intuition of $ T_1 \sim T_2 $ is that $ T_1 $ and $
%% T_2 $ have no conflicting type information. Two types are consistent
%% if one of them is $ \star $, or they have the same top-most type
%% constructor and the corresponding sub-parts are
%% consistent. Consistency is reflexive and symmetric, but not
%% transitive.

%% We write $ T_1 \smile T_2 $ when $ T_1 $ and $ T_2 $ are
%% shallowly-consistent, that is, if one of them is $ \star $, or they
%% have the same type constructor at the top. Shallow-inconsistency is
%% the root of all blame in lazy cast strategies -- casting a value to a
%% shallowly inconsistent type leads to a blame. Shallow-consistency is
%% reflexive, symmetric, but not transitive.

%% Let $ e $ range over terms. Most terms are standard for a simply-typed
%% lambda calculus. We annotate the codomain of a lambda abstraction
%% explicitly because we refer to it in the dynamic semantics. Becuase
%% this is a cast calculus, we have two additional terms: cast
%% expressions and blame expressions. Cast expressions perform runtime
%% type checks and blame expressions raise an error.

@@

The transition relation $\judgeCreduce{s}{s}$ is parameterized over
$applyCast$ to allow for the differences between D and UD.
%
When evaluating a cast expression, the machine moves the cast to the
continuation and evaluates the inner expression.
%% Other
%% transition rules starting from $ Eval $ states are standard.
To apply a casted function, the machine first casts $ v_1 $, the
operand, then applies the casted operand to $ v_2 $, the underlying
function, and finally cast the result of the function application.
%
To take out the first (resp. second) part of a casted pair, the
machine firstly take out the first (resp. second) part of $ v $, the
underlying pair, and cast the result.
%
In case splitting, if the target value is a left injection, the
machine moves to a state that will apply the first continuation
function to the value inside the left injection. The case for right
injection is similar.
%
When the target value is a casted sum, the machine moves the cast from
the target value to continuations functions.
%
To cast a value, the machine invokes $applyCast$ on the value.  If the
cast succeeds, the machine returns the result to the next
continuation.  If the cast fails, the machine halts with the blame
label.
%
%% When the continuation is $ stop $, the machine enters a halting
%% state. The When the machine is in a halting state, it stays in the
%% same state.

The reflexive transitive closure of reduction ($
\judgeCreduceTrans{s}{s} $) and evaluation ($ eval_\mathcal{C}(e) $)
are standard~\citep{felleisen03:_pllc}.

@@

        Shallow-consistency
	\fbox{$ T \smile T $}
	\begin{gather*}
	\inference{}{
		\star \smile \star
	} \quad
	\inference{}{
		\star \smile P
	} \quad
	\inference{}{
		P \smile \star
	} \\
	\inference{}{
		\iota \smile \iota
	} \quad
	\inference{}{
		T_{11} \rightarrow T_{12} \smile T_{21} \rightarrow T_{22}
	} \quad
	\inference{}{
		T_{11} \times T_{12} \smile T_{21} \times T_{22}
	} \quad
	\inference{}{
	T_{11} \plus T_1 \smile S_2 \plus T_2
	}
	\end{gather*}

	Value typing \fbox{$ v : T $}
	\begin{gather*}
	\dots \qquad
	\inference{
		v : I
	}{
		\vOOcast{v}{\cOOcast{I}{l}{\TOOdyn}} : \TOOdyn
	}
	\quad
	\inference{
		v : P_1 &
		P_1 \smile P_2
	}{
		\vOOcast{v}{\cOOcast{P_1}{l}{P_2}} : P_2
	}
	\end{gather*}
        \caption{Definition of machine state and auxilliary data
          structures.}
        \label{fig:state}
\end{figure}

Let $ v $ range over values. A value is either the unit, a function (a
closure), a pair, a left injection, a right injection, or a casted
value. As shown by the value typing rules, if the target type of a
casted value is the dynamic type, the underlying value must be of an
injectable type. The definition of injectable types depends on blame
strategies: for the \lazyD\ strategy, every pre-type is injectable;
for the \lazyUD\ strategy, a pre-types is injectable if and only if
all its sub-parts are the dynamic type. If the target type of a casted
value is a pre-type, the type of the underlying value must have the
same type constructor.

Let $ c $ range over casts. A cast is a triple of a type, a label, and a type.

Let $ o $ range over observations. They are what would be observed if
a program terminates. Observations include all value constructors and
blame.  The function converting values to observations ($ observe(v) =
o $) is defined in the natural way.

\begin{figure}
	%	Continuation typing \fbox{$ \kappa : T_1 \Longrightarrow T_2 $}
	%	\begin{gather*}
	%	\dots \quad
	%	\inference{
	%		c : T_1 \Longrightarrow T_2 &
	%		\kappa : T_2 \Longrightarrow T_3
	%	}{
	%		\langle c \rangle \kappa : T_1 \Longrightarrow T_3
	%	}
	%	\end{gather*}

@@

	Reduction \fbox{$ \judgeCreduce{s}{s} $}
	\[
	\begin{array}{rclr}
		\hiredrule{
		\sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\rho}{\kappa}
	}{
		\sOOinspect{e}{\rho}{
			\kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa}
%			\langle\cOOcast{T_1}{l}{T_2}\rangle\kappa
		}
	}{}
	\redrule{
		\sOOinspect{\eOOvar{x}}{\rho}{\kappa}
	}{	
		\sOOreturn{\rho(x)}{\kappa}
	}{}
	\redrule{
		\sOOinspect{\eOOsole}{\rho}{\kappa}
	}{
		\sOOreturn{\vOOtt}{\kappa}
	}{}
	\redrule{
		\sOOinspect{\eOOlam{T_1}{T_2}{x}{e}}{\rho}{\kappa}
	}{
		\sOOreturn{(\vOOfun{\rho}{x}{e})}{\kappa}
	}{}
	\redrule{
		\sOOinspect{(\eOOcons{e_1}{e_2})}{\rho}{\kappa}
	}{
		\sOOinspect{e_1}{\rho}{(\kOOconsI{e_2}{\rho}{\kappa})}
	}{}
	\redrule{
		\sOOinspect{(\eOOinl{e})}{\rho}{\kappa}
	}{
		\sOOinspect{e}{\rho}{(\kOOinl{\kappa})}
	}{}
	\redrule{
	\sOOinspect{(\eOOinr{e})}{\rho}{\kappa}
	}{
	\sOOinspect{e}{\rho}{(\kOOinr{\kappa})}
	}{}
	\redrule{
		\sOOinspect{(\eOOapp{e_1}{e_2})}{\rho}{\kappa}
	}{
\sOOinspect{e_1}{\rho}{(\kOOappI{e_2}{\rho}{\kappa})}}{}

@@

Let $ r $ range over cast results. A cast result is either a success, which 
brings back a value, or a failure, which brings a blame label.

Let $ s $ range over machine states. A state is either looking at an 
expression to decide what to do next, returning a value to a continuation, or 
halting with an observation.

Let $ \kappa $ range over continuations. $ \mathtt{stop} $ is the top 
continuation. The remaining continuations correspond to expressions. For 
example, $ (\mathtt{cons_1} \; e \; \rho \; \kappa) $ is the continuation where 
we are waiting for the value of the first argument to a $ \mathtt{cons} $. And 
the last continuation, $ \langle c \rangle \kappa $ is to cast the value before 
returning to $ \kappa $.

@@

	\end{array}
	\]
	\caption{Definition of $ applyCast $ for \lazyD}
	\label{fig:applyCast-D-C}
\end{figure}

@@

\figref{fig:normal-coercion} defines the normal coercion.
Types, pre-types, and blame labels are as before.
Let $ I $ range over injectable types. An injectable type is a type that can 
be cast directly to (injection) and from (projection) $ \TOOdyn $. The 
definition of injectable type depends on blame strategy. With \lazyD, all 
pre-types are injectable. With \lazyUD, only $ \POOfun{\TOOdyn}{\TOOdyn} $ 
is injectable. 
Let $ c $ range over coercions, a coercion is either injection, projection, 
identity, failure, sequencing, or function coercion.
Let $ \bar{c} $ range over wrapper coercions, which represent casts in $ 
\vOOcast{v}{c} $.
Let $ \hat{c} $ range over normal coercions. If we inline $ \bar{c} $ and 
re-order the cases, the definition of $ \hat{c} $ becomes:
\[
\begin{array}{rclr}
\stxrule{\hat{c}}{normal coercions}{
	\ncFun{\hat{c}}{\hat{c}} \mid
	\ncSeq{\ncProj{I}{l}}{\ncFun{\hat{c}}{\hat{c}}} \mid
	\ncSeq{\ncFun{\hat{c}}{\hat{c}}}{\ncInj{I}} \mid
	\ncSeq{\ncProj{I}{l}}{\ncSeq{\ncFun{\hat{c}}{\hat{c}}}{\ncInj{I}}} \mid
	\ncSeq{\ncProj{I}{l}}{\ncFail{l}}
}
\stxrulecont{
	\ncId \mid
	\ncProj{I}{l} \mid
	\ncInj{I} \mid
	\ncSeq{\ncProj{I}{l}}{\ncInj{I}} \mid
	\ncFail{l} \mid
}
\end{array}
\]

@@

	hypercoercion typing \fbox{$ c : T \Longrightarrow T $}
	\begin{gather*}
	\inference{}{\typingHC{\hyperCoercionI}{\TOOdyn}{\TOOdyn}}
	\quad
	\inference{
		\typingHC{h}{T_1}{P_1} &
		\typingHC{m}{P_1}{P_2} &
		\typingHC{t}{P_2}{T_2}
	}{
		\typingHC{\hyperCoercionC{h}{m}{t}}{T_1}{T_2}
	}
	\end{gather*}

	Head typing \fbox{$ \typingHC{h}{T}{P} $}
	\begin{gather*}
	\inference{}{\typingHC{\epsilon}{P}{P}}
	\quad
	\inference{}{\typingHC{?^l}{\TOOdyn}{I}}
	\end{gather*}

>>>>

\newcommand{\typingHC}[3]{#1 : #2 \Longrightarrow #3}
\newcommand{\hcvOOinj}[2]{\mathtt{inj} \; #2}
\newcommand{\hcvOOfun}[5]{\mathtt{fun} \; #1 \; #2 \; #3 \; #4 \; #5}
\newcommand{\hcvOOtt}[0]{\mathtt{unit}}
\newcommand{\hcvOOcons}[4]{\mathtt{cons}\;#1\;#2\;#3\;#4}
\newcommand{\hcvOOinl}[2]{\mathtt{inl}\;#1\;#2}
\newcommand{\hcvOOinr}[2]{\mathtt{inr}\;#1\;#2}
\newcommand{\hckOOmt}[0]{\mathtt{stop}}
\newcommand{\hckOOconsI}[3]{\mathtt{cons_1}\;#1\;#2\;#3}
\newcommand{\hckOOappII}[2]{\mathtt{app_2}\;#1\;#2}
\newcommand{\sidecond}[1]{\text{if}\;#1}
% Lazy D cast calculus on space-inefficient CEK
\newcommand{\judgeCreduce}[2]{#1 \longmapsto_{\mathcal{C}} #2}
\newcommand{\judgeCreduceTrans}[2]{#1 \longmapsto_{\mathcal{C}}^{*} #2}
\newcommand{\judgeCeval}[2]{eval_{\mathcal{C}}(#1) = #2}
\newcommand{\redrule}[3]{#1 & \longmapsto_\mathcal{C} & #2 & #3\\}
\newcommand{\hiredrule}[3]{\highlight{#1} & \highlight{\longmapsto_\mathcal{C}} & \highlight{#2} & \highlight{#3} \\}
% blame calculus on space-efficient CEK
\newcommand{\judgeSreduce}[3]{#2 \longmapsto_{\mathcal{S}(#1)} #3}
\newcommand{\judgeSreduceTrans}[3]{#2 \longmapsto_{\mathcal{S}(#1)}^{*} #3}
\newcommand{\judgeSeval}[3]{eval_{\mathcal{S}(#1)}(#2) = #3}
\newcommand{\redruleS}[3]{#1 & \longmapsto_{\mathcal{S}(C)} & #2 & #3\\}
\newcommand{\hiredruleS}[3]{\highlight{#1} & 
\highlight{\longmapsto_{\mathcal{S}(C)}} & \highlight{#2} & \highlight{#3} \\}
% Normal Coercion
\newcommand{\ncProj}[2]{#1?^{#2}}
\newcommand{\ncInj}[1]{#1!}
\newcommand{\ncId}[0]{\iota}
\newcommand{\ncSeq}[2]{#1;#2}
\newcommand{\ncFail}[1]{\bot^{#1}}
\newcommand{\ncFun}[2]{\POOfun{#1}{#2}}
\newcommand{\ncProd}[2]{\POOprod{#1}{#2}}
\newcommand{\ncSum}[2]{\POOsum{#1}{#2}}
% Hypercoercion
\newcommand{\hyperCoercionI}[0]{\mathtt{id\star}}
\newcommand{\hyperCoercionC}[3]{#1 \overset{#2}{\curvearrowright} #3}
% machine state simulations
\newcommand{\eqvS}[4]{#3 \approx_{\mathcal{S}\mathcal{S}} #4}
\newcommand{\eqvSD}[3]{#2 \approx_{\mathcal{SD}} #3}
% to-dos
\newcommand{\todoKC}[1]{\todo[inline]{KC needs to #1}}
\newcommand{\todoKCFixed}[0]{\todo[inline]{Fixed. -KC}}
% abbrev
\newcommand{\castCalculus}[0]{$\lambda_{\rightarrow}^{\langle\cdot\rangle}$}
% names
\newcommand{\ineffCEK}{$\mathcal{C}$}
\newcommand{\ineffCEKD}{$\mathcal{D}$}
\newcommand{\ineffCEKUD}{$\mathcal{UD}$}
\newcommand{\judgeDreduce}[2]{#1 \longmapsto_{\mathcal{D}} #2}
\newcommand{\judgeDreduceTrans}[2]{#1 \longmapsto_{\mathcal{D}}^{*} #2}
\newcommand{\judgeDeval}[2]{eval_{\mathcal{D}}(#1) = #2}
\newcommand{\judgeUDreduce}[2]{#1 \longmapsto_{\mathcal{UD}} #2}
\newcommand{\judgeUDreduceTrans}[2]{#1 \longmapsto_{\mathcal{UD}}^{*} #2}
\newcommand{\judgeUDeval}[2]{eval_{\mathcal{UD}}(#1) = #2}
\newcommand{\effCEK}[1]{$\mathcal{S}(#1)$}
\newcommand{\evalEqv}[2]{$eval_{\text{#1}} = eval_{\text{#2}}$}
\newcommand{\continue}[2]{continue(#2,#1)}
\newcommand{\highlight}[1]{{\color{red} #1}}

@@

	Consistency
	\fbox{$T_1 \sim T_2$}
	\begin{gather*}
	\inference{}{
		\star \sim \star
	} \quad
	\inference{}{
		\star \sim P
	} \quad
	\inference{}{
		P \sim \star
	} \\[1ex]
	\inference{}{
		\iota \sim \iota
	} \quad
	\inference{
		S_1 \sim S_2 &
		T_1 \sim T_2
	}{
		S_1 \rightarrow T_1 \sim S_2 \rightarrow T_2
	} \quad
	\inference{
		S_1 \sim S_2 &
		T_1 \sim T_2
	}{
		S_1 \times T_1 \sim S_2 \times T_2
	} \quad
	\inference{
		S_1 \sim S_2 &
		T_1 \sim T_2
	}{
		S_1 \plus T_1 \sim S_2 \plus T_2
	}
	\end{gather*}

	Term typing
	\fbox{$\judgetype{\Gamma}{e}{T}$}
	\begin{gather*}
          \dots \qquad
		\inference{
			\Gamma \vdash e : T_1 & T_1 \sim T_2
		}{
			\judgetype{\Gamma}{\eOOcast{e}{T_1}{l}{T_2}}{T_2}
		} \quad
		\inference{
		}{
			\judgetype{\Gamma}{\eOOblame{l}}{T}
		}
	\end{gather*}

@@

The syntax and static semantics is the same for the \lazyD{} and
\lazyUD{} cast calculi, and is reviewed in \figref{fig:blame-static}.
As usual, the important features are the cast expressions,
$\eOOcast{e}{T_1}{l}{T_2}$, which are responsible for runtime type
checking, and blame expressions, $\eOOblame{l}$, that raise errors.
The syntax and static semantics is the same that of
\citet{siek2009exploring} except for a few minor exceptions:
\begin{itemize}
\item We add sum, product, and unit types.
\item We separate types into those with a type constructor at the top,
  the \emph{pretypes} ($\POOunit$, functions, products, and sums),
  versus the dynamic type $\star$ (a.k.a. $\mathtt{Dyn}$ or
  $\mathbb{?}$).
\item We annotate the codomain of a lambda abstraction explicitly
  because we refer to it in the dynamic semantics. 
\end{itemize}
As usual, the source $T_1$ and target types $T_2$ of a cast
$\eOOcast{e}{T_1}{l}{T_2}$ are required to be consistent, written $T_1
\sim T_2$. The consistency relation is standard and defined in
\figref{fig:blame-static}.

%% defines the syntax of the cast calculus and
%% its static semantics. We extend the syntax of
%%  with sum and product types.  Let $T$ range
%% over types. A type is either the dynamic type$\star$(a.k.a. $
%% \mathtt{Dyn}$, $\mathbb{?}$, or$\mathtt{Unknown}$), or a type

%% As usual we write$T_1 \sim T_2$when$T_1$ and $T_2$ are
%% consistent. The intuition of $T_1 \sim T_2$ is that $T_1$and $
%% T_2$have no conflicting type information. Two types are consistent
%% if one of them is $\star$, or they have the same top-most type
%% constructor and the corresponding sub-parts are
%% consistent. Consistency is reflexive and symmetric, but not
%% transitive.

%% We write$T_1 \smile T_2$when$T_1$ and $T_2$ are
%% shallowly-consistent, that is, if one of them is $\star$, or they
%% have the same type constructor at the top. Shallow-inconsistency is
%% the root of all blame in lazy cast strategies -- casting a value to a
%% shallowly inconsistent type leads to a blame. Shallow-consistency is
%% reflexive, symmetric, but not transitive.

%% Let $e$ range over terms. Most terms are standard for a simply-typed
%% lambda calculus. We annotate the codomain of a lambda abstraction
%% explicitly because we refer to it in the dynamic semantics. Becuase
%% this is a cast calculus, we have two additional terms: cast
%% expressions and blame expressions. Cast expressions perform runtime
%% type checks and blame expressions raise an error.

@@

The transition relation $\judgeCreduce{s}{s}$ is parameterized over
$applyCast$ to allow for the differences between D and UD.
%
When evaluating a cast expression, the machine moves the cast to the
continuation and evaluates the inner expression.
%% Other
%% transition rules starting from$Eval$states are standard.
To apply a casted function, the machine first casts $v_1$, the
operand, then applies the casted operand to $v_2$, the underlying
function, and finally cast the result of the function application.
%
To take out the first (resp. second) part of a casted pair, the
machine firstly take out the first (resp. second) part of $v$, the
underlying pair, and cast the result.
%
In case splitting, if the target value is a left injection, the
machine moves to a state that will apply the first continuation
function to the value inside the left injection. The case for right
injection is similar.
%
When the target value is a casted sum, the machine moves the cast from
the target value to continuations functions.
%
To cast a value, the machine invokes $applyCast$ on the value.  If the
cast succeeds, the machine returns the result to the next
continuation.  If the cast fails, the machine halts with the blame
label.
%
%% When the continuation is $stop$, the machine enters a halting
%% state. The When the machine is in a halting state, it stays in the
%% same state.

The reflexive transitive closure of reduction ($\judgeCreduceTrans{s}{s}$) and 
evaluation ($eval_\mathcal{C}(e)$)
are standard~\citep{felleisen03:_pllc}.

@@

        Shallow-consistency
	\fbox{$T \smile T$}
	\begin{gather*}
	\inference{}{
		\star \smile \star
	} \quad
	\inference{}{
		\star \smile P
	} \quad
	\inference{}{
		P \smile \star
	} \\
	\inference{}{
		\iota \smile \iota
	} \quad
	\inference{}{
		T_{11} \rightarrow T_{12} \smile T_{21} \rightarrow T_{22}
	} \quad
	\inference{}{
		T_{11} \times T_{12} \smile T_{21} \times T_{22}
	} \quad
	\inference{}{
	T_{11} \plus T_1 \smile S_2 \plus T_2
	}
	\end{gather*}

	Value typing \fbox{$v : T$}
	\begin{gather*}
	\dots \qquad
	\inference{
		v : I
	}{
		\vOOcast{v}{\cOOcast{I}{l}{\TOOdyn}} : \TOOdyn
	}
	\quad
	\inference{
		v : P_1 &
		P_1 \smile P_2
	}{
		\vOOcast{v}{\cOOcast{P_1}{l}{P_2}} : P_2
	}
	\end{gather*}
        \caption{Definition of machine state and auxilliary data
          structures.}
        \label{fig:state}
\end{figure}

Let $v$ range over values. A value is either the unit, a function (a
closure), a pair, a left injection, a right injection, or a casted
value. As shown by the value typing rules, if the target type of a
casted value is the dynamic type, the underlying value must be of an
injectable type. The definition of injectable types depends on blame
strategies: for the \lazyD\ strategy, every pre-type is injectable;
for the \lazyUD\ strategy, a pre-types is injectable if and only if
all its sub-parts are the dynamic type. If the target type of a casted
value is a pre-type, the type of the underlying value must have the
same type constructor.

Let $c$ range over casts. A cast is a triple of a type, a label, and a type.

Let $o$ range over observations. They are what would be observed if
a program terminates. Observations include all value constructors and
blame.  The function converting values to observations ($observe(v) =
o$) is defined in the natural way.

\begin{figure}
	%	Continuation typing \fbox{$\kappa : T_1 \Longrightarrow T_2$}
	%	\begin{gather*}
	%	\dots \quad
	%	\inference{
	%		c : T_1 \Longrightarrow T_2 &
	%		\kappa : T_2 \Longrightarrow T_3
	%	}{
	%		\langle c \rangle \kappa : T_1 \Longrightarrow T_3
	%	}
	%	\end{gather*}

@@

	Reduction \fbox{$\judgeCreduce{s}{s}$}
	\[
	\begin{array}{rclr}
		\hiredrule{
		\sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\rho}{\kappa}
	}{
		\sOOinspect{e}{\rho}{
			\kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa}
%			\langle\cOOcast{T_1}{l}{T_2}\rangle\kappa
		}
	}{}
	\redrule{
		\sOOinspect{\eOOvar{x}}{\rho}{\kappa}
	}{	
		\sOOreturn{\rho(x)}{\kappa}
	}{}
	\redrule{
		\sOOinspect{\eOOsole}{\rho}{\kappa}
	}{
		\sOOreturn{\vOOtt}{\kappa}
	}{}
	\redrule{
		\sOOinspect{\eOOlam{T_1}{T_2}{x}{e}}{\rho}{\kappa}
	}{
		\sOOreturn{(\vOOfun{\rho}{x}{e})}{\kappa}
	}{}
	\redrule{
		\sOOinspect{(\eOOcons{e_1}{e_2})}{\rho}{\kappa}
	}{
		\sOOinspect{e_1}{\rho}{(\kOOconsI{e_2}{\rho}{\kappa})}
	}{}
	\redrule{
		\sOOinspect{(\eOOinl{e})}{\rho}{\kappa}
	}{
		\sOOinspect{e}{\rho}{(\kOOinl{\kappa})}
	}{}
	\redrule{
	\sOOinspect{(\eOOinr{e})}{\rho}{\kappa}
	}{
	\sOOinspect{e}{\rho}{(\kOOinr{\kappa})}
	}{}
	\redrule{
		\sOOinspect{(\eOOapp{e_1}{e_2})}{\rho}{\kappa}
	}{
\sOOinspect{e_1}{\rho}{(\kOOappI{e_2}{\rho}{\kappa})}}{}

@@

Let $r$ range over cast results. A cast result is either a success, which 
brings back a value, or a failure, which brings a blame label.

Let $s$ range over machine states. A state is either looking at an 
expression to decide what to do next, returning a value to a continuation, or 
halting with an observation.

Let $\kappa$ range over continuations. $\mathtt{stop}$ is the top 
continuation. The remaining continuations correspond to expressions. For 
example, $(\mathtt{cons_1} \; e \; \rho \; \kappa)$ is the continuation where 
we are waiting for the value of the first argument to a $\mathtt{cons}$. And 
the last continuation, $\langle c \rangle \kappa$ is to cast the value before 
returning to $\kappa$.

@@

	\end{array}
	\]
	\caption{Definition of $applyCast$for \lazyD}
	\label{fig:applyCast-D-C}
\end{figure}

@@

\figref{fig:normal-coercion} defines the normal coercion.
Types, pre-types, and blame labels are as before.
Let $I$ range over injectable types. An injectable type is a type that can 
be cast directly to (injection) and from (projection) $\TOOdyn$. The 
definition of injectable type depends on blame strategy. With \lazyD, all 
pre-types are injectable. With \lazyUD, only $\POOfun{\TOOdyn}{\TOOdyn}$
is injectable. 
Let $c$ range over coercions, a coercion is either injection, projection, 
identity, failure, sequencing, or function coercion.
Let $\bar{c}$ range over wrapper coercions, which represent casts in $
\vOOcast{v}{c}$.
Let $\hat{c}$ range over normal coercions. If we inline $\bar{c}$ and 
re-order the cases, the definition of $\hat{c}$ becomes:
\[
\begin{array}{rclr}
\stxrule{\hat{c}}{normal coercions}{
	\ncFun{\hat{c}}{\hat{c}} \mid
	\ncSeq{\ncProj{I}{l}}{\ncFun{\hat{c}}{\hat{c}}} \mid
	\ncSeq{\ncFun{\hat{c}}{\hat{c}}}{\ncInj{I}} \mid
	\ncSeq{\ncProj{I}{l}}{\ncSeq{\ncFun{\hat{c}}{\hat{c}}}{\ncInj{I}}} \mid
	\ncSeq{\ncProj{I}{l}}{\ncFail{l}}
}
\stxrulecont{
	\ncId \mid
	\ncProj{I}{l} \mid
	\ncInj{I} \mid
	\ncSeq{\ncProj{I}{l}}{\ncInj{I}} \mid
	\ncFail{l} \mid
}
\end{array}
\]

@@

	hypercoercion typing \fbox{$c : T \Longrightarrow T$}
	\begin{gather*}
	\inference{}{\typingHC{\hyperCoercionI}{\TOOdyn}{\TOOdyn}}
	\quad
	\inference{
		\typingHC{h}{T_1}{P_1} &
		\typingHC{m}{P_1}{P_2} &
		\typingHC{t}{P_2}{T_2}
	}{
		\typingHC{\hyperCoercionC{h}{m}{t}}{T_1}{T_2}
	}
	\end{gather*}

	Head typing \fbox{$\typingHC{h}{T}{P}$}
	\begin{gather*}
	\inference{}{\typingHC{\epsilon}{P}{P}}
	\quad
	\inference{}{\typingHC{?^l}{\TOOdyn}{I}}
	\end{gather*}

	Middle typing \fbox{$\typingHC{m}{T}{T}$}
	\begin{gather*}
	\inference{}{\typingHC{\POOunit}{\POOunit}{\POOunit}}
	\quad
	\inference{
		\typingHC{c_1}{T_3}{T_1} &
		\typingHC{c_2}{T_2}{T_4}
	}{
		\typingHC{\POOfun{c_1}{c_2}}{\POOfun{T_1}{T_2}}{\POOfun{T_3}{T_4}}
	}
	\\
	\inference{
		\typingHC{c_1}{T_1}{T_3} &
		\typingHC{c_2}{T_2}{T_4}
	}{
		\typingHC{\POOprod{c_1}{c_2}}{\POOprod{T_1}{T_2}}{\POOprod{T_3}{T_4}}
	}
	\quad
	\inference{
		\typingHC{c_1}{T_1}{T_3} &
		\typingHC{c_2}{T_2}{T_4}
	}{
		\typingHC{\POOsum{c_1}{c_2}}{\POOsum{T_1}{T_2}}{\POOsum{T_3}{T_4}}
	}
		\end{gather*}
